The dynamics of starvation and recovery

### 2017-06-09 Manuscript/Nature-Submission/14.tex
Nature submission

\noindent {\bf Nutritional state-structured model (NSM)}\\
We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$. The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity $C$. The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes. In the supplement we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we have described below (please see Supporting Information for a detailed derivation).

Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$  (see Supporting Information for detailed a description and derivation of all rates).  Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource. %and $\approx 2$.
Resources are eaten by the consumers at rate $\rho R + \delta$, which accounts for somatic growth and maintenance for starving consumers, respectively, and at rate $\beta$ by full consumers which is a constant representing maximal maintenance and somatic growth.
%This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
The NSM represents a fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\cite{Persson:1998hz}.

@@

Nearly all of the rates described in the NSM are determined by consumer
metabolism, which can be used to describe a variety of organismal features
\cite{Brown:2004wq}.  The scaling relation between an organism's metabolic
rate $B$ and its body mass $M$ at reproductive maturity is known to scale as
$B = B_0 M^\eta$~\cite{West:2002it}, where the scaling exponent $\eta$ is
typically close to $2/3$ or $3/4$ for metazoans (e.g., \cite{Brown:2004wq}),
and has taxonomic shifts for unicellular species between $\eta\approx 1$ in
eukaryotes and $\eta\approx 1.76$ in bacteria
\cite{DeLong:2010dy,Kempes:2012hy}.
%Justin adding an explanation for beta

@@

%Minimal mass bound
%A theoretical lower bound on mammalian body size occurs when $\epsilon=0$ where mammals have no metabolic reserves and immediately starve; this occurs at a size of $M=0$, which is obviously infeasible.
%Such prediction gives an extreme limit on size but does not account for the subtleties of starvation dynamics
%that may limit body size (among other factors).
The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis \cite{allen2002,enquist1998}.
The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size (e.g. \cite{allen2002,enquist1998}). %Previous studies have focused on t
This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism (e.g. $N^*\propto M^{-3/4}/B_{0}$)  (e.g. \cite{allen2002,enquist1998}).
This is usually stated as $B_{\rm tot}=N^*B(M)=Q$ where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{allen2002,enquist1998}.
Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
This result if remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to the energy equivalence result.
Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this is a hard limit where deviations outside of this range are energetically suboptimal. %, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, occurring at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}). This asymptotic behavior represents an upper bound on mammalian body size and occurs at $M_{\rm max}=6.5\times10^7$.
Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
We note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}. % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here

@@

\begin{figure*}
\centering
\includegraphics[width=0.75\textwidth]{fig_Invasion.pdf}
\caption{ Invasion feasibility for organisms with a proportional change in
  mass $\chi$ against a population with a resident body mass $M$.  The blue
  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
  red region denotes values of $\chi$ that result in a mass that is below the
  starvation threshold and is thus infeasible.
  Arrows point to the predicted optimal mass $M_{\rm opt}=1.73\times 10^7$, which serves as an evolutionary attractor for body mass.
  The black wedge points to the largest body mass known for terrestrial mammals at $7.74\times10^7$g~\cite{Smith:2010p3442}.}
\label{fig:invasion}
\end{figure*}

>>>>

\noindent {\bf Nutritional state-structured model (NSM)}\\
We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity $c$, which we non-dimensionalize to equal one.
Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$.  Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ converts resources to somatic growth of consumer tissue. %and $\approx 2$.
Resources are eaten by the consumers at rate $\rho R + \delta$, which accounts for somatic growth and maintenance for starving consumers, respectively, and at rate $\beta$ by full consumers.
%This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
The NSM represents a fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\cite{Persson:1998hz}.

@@

Nearly all of the rates described in the NSM are determined by consumer
metabolism, which can be used to describe a variety of organismal features
\cite{Brown:2004wq}.  The scaling relation between an organism's metabolic
rate $B$ and its body mass $M$ at reproductive maturity is known to scale as
$B = B_0 M^\eta$~\cite{West:2002it}, where the scaling exponent $\eta$ is
typically close to $2/3$ or $3/4$ for metazoans (e.g., \cite{Brown:2004wq}),
and has taxonomic shifts for unicellular species between $\eta\approx 1$ in
eukaryotes and $\eta\approx 1.76$ in bacteria
\cite{DeLong:2010dy,Kempes:2012hy}.
%Justin adding an explanation for beta
{\color{blue}An organism's metabolic rate $B$ is proportional to the cost of tissue
maintenance in the absence of growth (i.e., when the body mass is $M$).  By
definition $B=\beta/\xi$, where $\beta$ is the rate at which resources are
consumed for full consumers (see Eq.~\eqref{eq:system}) and where $\xi$ is
related to the conversion efficiency of resource to consumer tissue
(Supporting Information).
}

@@

%Minimal mass bound
%A theoretical lower bound on mammalian body size occurs when $\epsilon=0$ where mammals have no metabolic reserves and immediately starve; this occurs at a size of $M=0$, which is obviously infeasible.
%Such prediction gives an extreme limit on size but does not account for the subtleties of starvation dynamics
%that may limit body size (among other factors).
The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis \cite{allen2002,enquist1998}.
The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size (e.g. \cite{allen2002,enquist1998}). %Previous studies have focused on t
This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism (e.g. $N^*\propto M^{-3/4}/B_{0}$)  (e.g. \cite{allen2002,enquist1998}).
This is usually stated as $B_{\rm tot}=N^*B(M)=Q$ where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{allen2002,enquist1998}.
Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
This result if remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to the energy equivalence result.
Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this is a hard limit where deviations outside of this range are energetically suboptimal. %, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, occurring at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}). This asymptotic behavior represents an upper bound on mammalian body size and occurs at $M_{\rm max}=6.5\times10^7$.
Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.75\textwidth]{fig_Invasion.pdf}
\caption{ Invasion feasibility for organisms with a proportional change in
  mass $\chi$ against a population with a resident body mass $M$.  The blue
  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
  red region denotes values of $\chi$ that result in a mass that is below the
  starvation threshold and is thus infeasible.
  Arrows point to the predicted optimal mass $M_{\rm opt}=1.73\times 10^7$, which serves as an evolutionary attractor for body mass.
  The black wedge points to the largest body mass known for terrestrial mammals at $7.4\times10^7$g~\cite{Smith:2010p3442}.}
\label{fig:invasion}
\end{figure*}

### 2017-06-09 Manuscript/Nature-Submission/18.tex
Nature submission

The time to death should follow a similar relation, but defined by a lower
fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$ where $\epsilon_\mu < \epsilon_\sigma$.
Suppose, for example, that an organism dies once it has digested all fat and
muscle tissues, and that muscle tissue scales with body mass according to
$M_{\rm musc}=u_{0}M^{\zeta}$.  This gives
$\epsilon_{\mu}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
mass has been shown to be roughly proportional to body mass~\cite{Folland:2008ij} in
mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant. The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or
\begin{eqnarray}
\label{mutimescale}
t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
\end{eqnarray}
and $\mu=1/t_{\mu}$.

@@

Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal. %, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}). This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero causing the timescales to become infinite (see Equation \ref{mutimescale}) and the rates $\mu$ and $\lambda$ to equal zero.
The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$. This point predicts a strong upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}. % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here

@@

@@

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
\caption{\small{ Invasion feasibility for organisms with a proportional change in
  mass $\chi$ against a population with a resident body mass $M$.  The blue
  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
  red region denotes values of $\chi$ that result in a mass that is below the
  starvation threshold and are thus infeasible.
  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}\label{fig:invasion}}
\end{figure}

@@

\bibliography{aa_starving3}

>>>>

The time to death should follow a similar relation, but defined by a lower
fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$ where $\epsilon_\mu < \epsilon_\sigma$.
Suppose, for example, that an organism dies once it has digested all fat and
muscle tissues, and that muscle tissue scales with body mass according to
$M_{\rm musc}=u_{0}M^{\zeta}$.  This gives
$\epsilon_{\mu}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
mass has been shown to be roughly proportional to body mass~\cite{Folland:2008ij} in
mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant. The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or $t_{\mu}=-M^{1-\eta}\ln\left(\epsilon_{\mu}\right)/a^{\prime}-t_{\sigma}$,
% \begin{eqnarray}
% \label{mutimescale}
% t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
% \end{eqnarray}
and $\mu=1/t_{\mu}$.

@@

Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal. %, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}). This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$. This point predicts a strong upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}. % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
\caption{\small{ Invasion feasibility for organisms with a proportional change in
  mass $\chi$ against a population with a resident body mass $M$.  The blue
  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
  red region denotes values of $\chi$ that result in a mass that is below the
  starvation threshold and are thus infeasible.
  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}\label{fig:invasion}}
\end{figure}

@@

@@

% \bibliography{aa_starving3}
\begin{thebibliography}{10}

\bibitem{Martin:1987dl}
Martin TE
\newblock (1987) {Food as a limit on breeding birds: A life-history
  perspective}.
\newblock \emph{Annu. Rev. Ecol. Syst.} 18:453--487.

\bibitem{Kirk:1997cc}
Kirk KL
\newblock (1997) {Life-history responses to variable environments: Starvation
  and reproduction in planktonic rotifers}.
\newblock \emph{Ecology} 78:434--441.

\bibitem{Kempes:2012hy}
Kempes CP, Dutkiewicz S, Follows MJ
\newblock (2012) {Growth, metabolic partitioning, and the size of
  microorganisms}.
\newblock \emph{Proc. Natl. Acad. Sci. USA} 109:495--500.

\bibitem{Mangel:1988uaa}
Mangel M, Clark CW
\newblock (1988) \emph{{Dynamic Modeling in Behavioral Ecology}}
\newblock (Princeton University Press, Princeton).

\bibitem{Mangel:2014kz}
Mangel M
\newblock (2014) {Stochastic dynamic programming illuminates the link between
  environment, physiology, and evolution}.
\newblock \emph{B. Math. Biol.} 77:857--877.

\bibitem{Yeakel:2013hi}
Yeakel JD, Dominy NJ, Koch PL, Mangel M
\newblock (2014) {Functional morphology, stable isotopes, and human evolution:
  a model of consilience}.
\newblock \emph{Evolution} 68:190--203.

\bibitem{Morris:1987eo}
Morris DW
\newblock (1987) {Optimal allocation of parental investment}.
\newblock \emph{Oikos} 49:332--339.

\bibitem{Tveraa:2003fq}
Tveraa T, Fauchald P, Henaug C, Yoccoz NG
\newblock (2003) {An examination of a compensatory relationship between food
  limitation and predation in semi-domestic reindeer}.
\newblock \emph{Oecologia} 137:370--376.

\bibitem{Daan:1988va}
Daan S, Dijkstra C, Drent R, Meijer T
\newblock (1989) in \emph{Acta XIX Congressus Internationalis Ornithologici,
  Volume I: Proceedings XIX International Ornithological Congress, 1986,
  Ottawa}, ed{} Ouellet H
\newblock (Proceedings XIX International Ornithological Congress, Ottawa), pp
  392--407.

\bibitem{Jacot:2009dv}
Jacot A, Valcu M, van Oers K, Kempenaers B
\newblock (2009) {Experimental nest site limitation affects reproductive
  strategies and parental investment in a hole-nesting passerine}.
\newblock \emph{Animal Behaviour} 77:1075--1083.

\bibitem{Stearns:1989ip}
Stearns SC
\newblock (1989) {Trade-offs in life-history evolution}.
\newblock \emph{Funct. Ecol.} 3:259.

\bibitem{Barboza:2002in}
Barboza P, Jorde D
\newblock (2002) {Intermittent fasting during winter and spring affects body
  composition and reproduction of a migratory duck}.
\newblock \emph{J Comp Physiol B} 172:419--434.

\bibitem{Threlkeld:1976ih}
Threlkeld ST
\newblock (1976) {Starvation and the size structure of zooplankton
  communities}.
\newblock \emph{Freshwater Biol.} 6:489--496.

\bibitem{Weber:1998jg}
Weber TP, Ens BJ, Houston AI
\newblock (1998) {Optimal avian migration: A dynamic model of fuel stores and
  site use}.
\newblock \emph{Evolutionary Ecology} 12:377--401.

\bibitem{Mduma:1999cp}
Mduma SAR, Sinclair ARE, Hilborn R
\newblock (1999) {Food regulates the Serengeti wildebeest: a 40-year record}.
\newblock \emph{J. Anim. Ecol.} 68:1101--1122.

\bibitem{Moore:2014hi}
Moore JW, Yeakel JD, Peard D, Lough J, Beere M
\newblock (2014) {Life-history diversity and its importance to population
  stability and persistence of a migratory fish: steelhead in two large North
  American watersheds}.
\newblock \emph{J. Anim. Ecol.} 83:1035--1046.

\bibitem{Mead:1989dt}
Mead RA
\newblock (1989) in \emph{Carnivore Behavior, Ecology, and Evolution}, ed{}
  Gittleman JL
\newblock (Springer US, Ithaca), pp 437--464.

\bibitem{Sandell:1990kw}
Sandell M
\newblock (1990) {The evolution of seasonal delayed implantation}.
\newblock \emph{Q Rev Biol} 65:23--42.

\bibitem{Bulik:1999eo}
Bulik CM, {et~al.}
\newblock (1999) {Fertility and reproduction in women with anorexia nervosa}.
\newblock \emph{J. Clin. Psychiat.} 60:130--135.

\bibitem{Trites:2003cc}
Trites AW, Donnelly CP
\newblock (2003) {The decline of Steller sea lions Eumetopias jubatus in
  Alaska: a review of the nutritional stress hypothesis}.
\newblock \emph{Mammal Rev.} 33:3--28.

\bibitem{Glazier:2009hq}
Glazier DS
\newblock (2009) {Metabolic level and size scaling of rates of respiration and
  growth in unicellular organisms}.
\newblock \emph{Funct. Ecol.} 23:963--968.

\bibitem{Kooi2000}
Kooijman SALM
\newblock (2000) \emph{{Dynamic Energy and Mass Budgets in Biological Systems}}
\newblock (Cambridge).

\bibitem{Sousa:2010ez}
Sousa T, Domingos T, Poggiale JC, Kooijman SALM
\newblock (2010) {Dynamic energy budget theory restores coherence in biology}.
\newblock \emph{Philos. T. Roy. Soc. B} 365:3413--3428.

\bibitem{Diekmann:2010da}
Diekmann O, Metz JAJ
\newblock (2010) {How to lift a model for individual behaviour to the
  population level?}
\newblock \emph{Philos. T. Roy. Soc. B} 365:3523--3530.

\bibitem{murdoch:2003}
Murdoch WW, Briggs CJ, Nisbet RM
\newblock (2003) \emph{{Consumer-resource Dynamics}}, Monographs in population
  biology
\newblock (Princeton University Press, Princeton) Vol.{}~36.

\bibitem{Benichou:2014wu}
Benichou O, Redner S
\newblock (2014) {Depletion-Controlled Starvation of a Diffusing Forager}.
\newblock \emph{arXiv} pp 1--5.

\bibitem{Benichou:2016wl}
B{\'e}nichou O, Chupeau M, Redner S
\newblock (2016) {Role of depletion on the dynamics of a diffusing forager}.
\newblock \emph{Journal of Physics A: {\ldots}}.

\bibitem{Chupeau:2016jf}
Chupeau M, B{\'e}nichou O, Redner S
\newblock (2016) {Universality classes of foraging with resource renewal}.
\newblock \emph{Phys. Rev. E} 93:032403.

\bibitem{Persson:1998hz}
Persson L, Leonardsson K, De~Roos AM, Gyllenberg M, Christensen B
\newblock (1998) {Ontogenetic scaling of foraging rates and the dynamics of a
  size-structured consumer-resource model}.
\newblock \emph{Theor Popul Biol} 54:270--293.

\bibitem{murray2011mathematical}
Murray JD
\newblock (2011) \emph{{Mathematical Biology: I. An Introduction}},
  Interdisciplinary Applied Mathematics
\newblock (Springer New York) Vol.{} 110.

\bibitem{Strogatz:2008wo}
Strogatz SH
\newblock (2008) \emph{{Nonlinear Dynamics and Chaos: With Applications to
  Physics, Biology, Chemistry, and Engineering}}, Studies in nonlinearity
\newblock (Westview Press, Boulder).

\bibitem{GuckHolmes}
Guckenheimer J, Holmes P
\newblock (1983) \emph{{Nonlinear Oscillations, Dynamical Systems, and
  Bifurcations of Vector Fields}}
\newblock (Springer, New York).

\bibitem{Gross:2004p2428}
Gross T, Feudel U
\newblock (2004) {Analytical search for bifurcation surfaces in parameter
  space}.
\newblock \emph{Physica D} 195:292--302.

\bibitem{Hastings:2001jh}
Hastings A
\newblock (2001) {Transient dynamics and persistence of ecological systems}.
\newblock \emph{Ecol. Lett.} 4:215--220.

\bibitem{Neubert:1997wk}
Neubert M, Caswell H
\newblock (1997) {Alternatives to resilience for measuring the responses of
  ecological systems to perturbations}.
\newblock \emph{Ecology} 78:653--665.

\bibitem{Caswell:2005eo}
Caswell H, Neubert MG
\newblock (2005) {Reactivity and transient dynamics of discrete-time ecological
  systems}.
\newblock \emph{J Differ Equ Appl} 11:295--310.

\bibitem{Neubert:2009td}
Neubert M, Caswell H
\newblock (2009) {Detecting reactivity}.
\newblock \emph{Ecology} 90:2683--2688.

\bibitem{Yodzis:1992hg}
Yodzis P, Innes S
\newblock (1992) {Body size and consumer-resource dynamics}.
\newblock \emph{Am. Nat.} 139:1151--1175.

\bibitem{Brown:2004wq}
Brown J, Gillooly J, Allen A, Savage V, West G
\newblock (2004) {Toward a metabolic theory of ecology}.
\newblock \emph{Ecology} 85:1771--1789.

\bibitem{West:2002it}
West GB, Woodruff WH, Brown JH
\newblock (2002) {Allometric scaling of metabolic rate from molecules and
  mitochondria to cells and mammals.}
\newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.

\bibitem{DeLong:2010dy}
DeLong JP, Okie JG, Moses ME, Sibly RM, Brown JH
\newblock (2010) {Shifts in metabolic scaling, production, and efficiency
  across major evolutionary transitions of life}.
\newblock \emph{Proc. Natl. Acad. Sci. USA} 107:12941--12945.

\bibitem{West:2001bv}
West GB, Brown JH, Enquist BJ
\newblock (2001) {A general model for ontogenetic growth}.
\newblock \emph{Nature} 413:628--631.

\bibitem{moses2008rmo}
Moses ME, {et~al.}
\newblock (2008) {Revisiting a Model of Ontogenetic Growth: Estimating Model
  Parameters from Theory and Data.}
\newblock \emph{Am. Nat.} 171:632--645.

\bibitem{gillooly2002esa}
Gillooly JF, Charnov EL, West GB, Savage VM, Brown JH
\newblock (2002) {Effects of size and temperature on developmental time}.
\newblock \emph{Nature} 417:70--73.

\bibitem{hou}
Hou C, {et~al.}
\newblock (2008) {Energy uptake and allocation during ontogeny}.
\newblock \emph{Science} 322:736--739.

\bibitem{bettencourt}
Bettencourt LMA, Lobo J, Helbing D, Kuhnert C, West GB
\newblock (2007) {Growth, innovation, scaling, and the pace of life in cities}.
\newblock \emph{Proc. Natl. Acad. Sci. USA} 104:7301--7306.

\bibitem{Folland:2008ij}
Folland JP, Mc~Cauley TM, Williams AG
\newblock (2008) {Allometric scaling of strength measurements to body size}.
\newblock \emph{Eur J Appl Physiol} 102:739--745.

\bibitem{Brown:1993p708}
Brown J, Marquet P, Taper M
\newblock (1993) {Evolution of body size: consequences of an energetic
  definition of fitness}.
\newblock \emph{Am. Nat.} 142:573--584.

\bibitem{West:1997cg}
West GB, Brown JH, Enquist BJ
\newblock (1997) {A general model for the origin of allometric scaling laws in
  biology}.
\newblock \emph{Science} 276:122--126.

\bibitem{West:2002ud}
West GB, Woodruff WH, Brown JH
\newblock (2002) {Allometric scaling of metabolic rate from molecules and
  mitochondria to cells and mammals.}
\newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.

\bibitem{Millar:1990p923}
Millar J, Hickling G
\newblock (1990) {Fasting endurance and the evolution of mammalian body size}.
\newblock \emph{Funct. Ecol.} 4:5--12.

\bibitem{Alroy:1998p1594}
Alroy J
\newblock (1998) {Cope's rule and the dynamics of body mass evolution in North
  American fossil mammals}.
\newblock \emph{Science} 280:731--734.

\bibitem{Clauset:2009fh}
Clauset A, Redner S
\newblock (2009) {Evolutionary model of species body mass diversification}.
\newblock \emph{Phys. Rev. Lett.} 102:038103.

\bibitem{Smith:2010p3442}
Smith FA, {et~al.}
\newblock (2010) {The evolution of maximum body size of terrestrial mammals}.
\newblock \emph{Science} 330:1216--1219.

\bibitem{Saarinen:2014br}
Saarinen JJ, {et~al.}
\newblock (2014) {Patterns of maximum body size evolution in Cenozoic land
  mammals: Eco-evolutionary processes and abiotic forcing}.
\newblock \emph{Proc Biol Sci} 281:20132049.

\bibitem{Damuth:1987kr}
Damuth J
\newblock (1987) {Interspecific allometry of population density in mammals and
  other animals: the independence of body mass and population energy-use}.
\newblock \emph{Biol. J. Linn. Soc.} 31:193--246.

\bibitem{DeLong:2012kw}
DeLong JP, Vasseur DA
\newblock (2012) {A dynamic explanation of size{\textendash}density scaling in
  carnivores}.
\newblock \emph{Ecology} 93:470--476.

\bibitem{allen2002}
Allen AP, Brown JH, Gillooly JF
\newblock (2002) {Global biodiversity, biochemical kinetics, and the
  energetic-equivalence rule}.
\newblock \emph{Science} 297:1545--1548.

\bibitem{enquist1998}
Enquist BJ, Brown JH, West GB
\newblock (1998) {Allometric scaling of plant energetics and population
  density}.
\newblock \emph{Nature} 395:163--165.

\bibitem{Kempes:2016}
Kempes CP, Wang L, Amend JP, Doyle J, Hoehler T
\newblock (2016) {Evolutionary tradeoffs in cellular composition across diverse
  bacteria}.
\newblock \emph{ISME J} 10:2145--2157.

\bibitem{tilman1981}
Tilman D
\newblock (1981) {Tests of resource competition theory using four species of
  lake michigan algae}.
\newblock \emph{Ecology} 62:802--815.

\bibitem{dutkiewicz2009}
Dutkiewicz S, Follows MJ, Bragg JG
\newblock (2009) {Modeling the coupling of ocean ecology and biogeochemistry}.
\newblock \emph{Global Biogeochem. Cycles} 23:1--15.

\bibitem{barton2010}
Barton AD, Dutkiewicz S, Flierl G, Bragg J, Follows MJ
\newblock (2010) {Patterns of diversity in marine phytoplankton}.
\newblock \emph{Science} 327:1509--1511.

\end{thebibliography}
